---
title: "Data_cleaning_and_rate_estimations"
output: html_document
date: '2025-09-19'
---

#Install package in case they are not in installed
```{r}
rm(list=ls())

if ("devtools" %in% rownames(installed.packages()) == 'FALSE') install.packages('devtools') 
library(devtools)
if ("segmented" %in% rownames(installed.packages()) == 'FALSE') install.packages('segmented') 
if ("plotrix" %in% rownames(installed.packages()) == 'FALSE') install.packages('plotrix') 
if ("gridExtra" %in% rownames(installed.packages()) == 'FALSE') install.packages('gridExtra') 
if ("LoLinR" %in% rownames(installed.packages()) == 'FALSE') install_github('colin-olito/LoLinR') 
if ("lubridate" %in% rownames(installed.packages()) == 'FALSE') install.packages('lubridate') 
if ("chron" %in% rownames(installed.packages()) == 'FALSE') install.packages('chron') 
if ("plyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('plyr') 
if ("dplyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('dplyr') 
```

#Load necessary package
```{r}
library("ggplot2")
library("segmented")
library("plotrix")
library("gridExtra")
library("LoLinR")
library("lubridate")
library("chron")
library('plyr')
library('dplyr')
library('tidyverse')
library('dplyr')
library('tidyr')
library(readxl)
library(stringr)
library(lattice)
```

#######Data cleaning & rate estimations######

#Make a list of the respirometry files
```{r}
# bring in the oxygen files

file.names<-basename(list.files(path = path.p, pattern = "csv$", recursive = TRUE)) #list all csv file names in the folder and subfolders
file.names

#basename above removes the subdirectory name from the file

file.names<-basename(list.files(path = path.p, pattern = "csv$", recursive = TRUE))
file.names.full<-list.files(path = path.p, pattern = "csv$", recursive = TRUE) #list all csv file names in the folder and subfolders

file.names.full
```

#generate an empty 3 column dataframe with specific column names
```{r}

#generate a 3 column dataframe with specific column names
Photo.R<- data.frame(matrix(NA, nrow=length(file.names), ncol=4))
colnames(Photo.R) <- c("ID","Intercept", "umol.L.sec","Temp.C")
View(Photo.R)

Photo.R<-read.csv(file=paste0(path.p,"/../Photo.R.csv"), header = T) 
```

#Loading the file for metadata
```{r}

#Load Sample Info/your respiration data file, with all the times, etc.
#Sample.Info <- read.csv(file=paste0(path.p,"/../Panama MetaData/Nubbin_Sample_Info_T0_Panama_QC.csv"), header=T) #read sample.info data
#Sample.Info <- read.csv(file=paste0(path.p,"/../metadata.csv"), header=T) #read sample.info data
Sample.Info <- read.csv(file=paste0(path.p,"/../Cni_metadata.csv"), header=T) 
View(Sample.Info)
```

# make start and stop times real times
```{r}

Sample.Info$Start.time <- as.POSIXct(Sample.Info$Start.time,format="%H:%M:%S", tz = "") #convert time from character to time
Sample.Info$Stop.Time <- as.POSIXct(Sample.Info$Stop.Time,format="%H:%M:%S", tz = "") #convert time from character to time
#Add names for respiration for for loop 
#PR<-c('Photo','Resp')
```

#Calculating O2 consumption for every file in list
```{r}

for(i in 1:length(file.names.full)) { # for every file in list calculate O2 evolution and add the data to the Photo.R dataframe

  #find the lines in sample info that have the same file name that is being brought in
  FRow<-which(Sample.Info$ID==strsplit(file.names[i],'.csv'))

  # read in the O2 data one by one
  Photo.Data1 <-read.csv(file.path(path.p,file.names.full[i]), skip = 1, header=T) # skips the first line
  Photo.Data1  <- Photo.Data1[,c("Time","Value","Temp")] #subset columns of interest
  Photo.Data1$Time <- as.POSIXct(Photo.Data1$Time,format="%H:%M:%S", tz = "") #convert time from character to time
  Photo.Data1 <- na.omit(Photo.Data1) #omit NA from data frame



# clean up some of the data
    n<-dim(Photo.Data1)[1] # length of full data
    Photo.Data1 <-Photo.Data1[15:(n-3),] #start at data point ~15s in to avoid excess noise from start of run and remove last 3 lines containing text

    n<-dim(Photo.Data1)[1] #list length of trimmed data
    Photo.Data1$sec <- 1:n #set seconds by one from start to finish of run in a new column
    #Save plot prior to and after data thinning to make sure thinning is not too extreme
    rename <- sub(".csv","", file.names[i]) # remove all the extra stuff in the file name

  pdf(paste0("Output/",rename,"thinned.pdf")) # open the graphics device

    par(omi=rep(0.3, 4)) #set size of the outer margins in inches
    par(mfrow=c(1,2)) #set number of rows and columns in multi plot graphic

    plot(Value ~ sec, data=Photo.Data1 , xlab='Time (seconds)', ylab=expression(paste(' O'[2],' (',mu,'mol/L)')), type='n', axes=FALSE) #plot (empty plot to fill) data as a function of time

    usr  <-  par('usr') # extract the size of the figure margins

    rect(usr[1], usr[3], usr[2], usr[4], col='grey90', border=NA) # put a grey background on the plot

    whiteGrid() # make a grid
    box() # add a box around the plot

    points(Photo.Data1 $Value ~ Photo.Data1 $sec, pch=16, col=transparentColor('dodgerblue2', 0.6), cex=1.1)

    axis(1) # add the x axis
    axis(2, las=1) # add the y-axis

# Thin the data to make the code run faster
    Photo.Data.orig<-Photo.Data1 #save original unthinned data
    Photo.Data1 <-  thinData(Photo.Data1 , by=15)$newData1 #thin data by every 15 points for all the O2 values

    Photo.Data1$sec <- as.numeric(rownames(Photo.Data1 )) #maintain numeric values for time

    Photo.Data1$Temp<-NA # add a new column to fill with the thinned data

    Photo.Data1$Temp <-  thinData(Photo.Data.orig, xy = c(1,3), by=15)$newData1[,2] #thin data by every 15 points for the temp values

#plot the thinned data

  plot(Value ~ sec, data=Photo.Data1 , xlab='Time (seconds)', ylab=expression(paste(' O'[2],' (',mu,'mol/L)')), type='n', axes=FALSE)
#plot thinned data
  usr  <-  par('usr')
  rect(usr[1], usr[3], usr[2], usr[4], col='grey90', border=NA)
  whiteGrid()
  box()
  points(Photo.Data1 $Value ~ Photo.Data1 $sec, pch=16, col=transparentColor('dodgerblue2', 0.6), cex=1.1)
  axis(1)
  axis(2, las=1)
    ##Olito et al. 2017: It is running a bootstrapping technique and calculating the rate based on density
    #option to add multiple outputs method= c("z", "eq", "pc")
    Regs  <-  rankLocReg(xall=Photo.Data1$sec, yall=Photo.Data1$Value, alpha=0.5, method="pc", verbose=TRUE)
    # add the regression data
    plot(Regs)
    dev.off()

    # fill in all the O2 consumption and rate data
    Photo.R[i,2:3] <- Regs$allRegs[1,c(4,5)] #inserts slope and intercept in the dataframe
    Photo.R[i,1] <- rename #stores the file name in the Date column
    Photo.R[i,4] <- mean(Photo.Data1$Temp, na.rm=T)  #stores the Temperature in the Temp.C column
    #Photo.R[i,5] <- PR[j] #stores whether it is photosynthesis or respiration

    # rewrite the file every time... I know this is slow, but it will save the data that is already run
}

write.csv(Photo.R, 'Output/Photo.R.csv')
```

#read in Photo.R file so you don't need to run entire for loop again
```{r}
Photo.R <- read.csv('Output/Photo.R.csv')
Photo.R <- na.omit(Photo.R)

Photo.R$ID <- gsub( '.{3}$', '', Photo.R$ID)

#Photo.R$ID <- substr(Photo.R$ID, 1 , nchar(Photo.R$ID - 3))
View(Photo.R)
```

#Joint metadata with Photo.R
```{r}
Photo.R_full<-left_join(Photo.R, Sample.Info, by = "ID")
View(Photo.R_full) 
```

# Calculate R rate
```{r}
#Convert sample volume to mL
Photo.R_full$Volume <- Photo.R_full$Volume/1000 #calculate volume

#Account for chamber volume to convert from umol L-1 s-1 to umol s-1. This standardizes across water volumes (different because of algae size) and removes per Liter 
Photo.R_full$umol.sec <- Photo.R_full$umol.L.sec*Photo.R_full$Volume

View(Photo.R_full)
```

#Account for blank rate by temperature
```{r}
#convert character columns to factors

Photo.R_full <- Photo.R_full %>%
  mutate_if(sapply(., is.character), as.factor)

# make the blank column a factor
Photo.R_full$Blank<-as.factor(Photo.R_full$Blank)
View(Photo.R_full)

write.csv(Photo.R_full, 'Output/Photo.R_full.csv') 
```

#aggregate certain columns together for easier viewing
```{r}
photo.blnk <- aggregate(umol.sec ~ ID*Temp.Cat*Light_Dark*Run*Blank, data=Photo.R_full, mean)
```

# pull out only the blanks
```{r}
#photo.blnk<-photo.blnk[photo.blnk$Species=='BK',]
photo.blnk<-photo.blnk[photo.blnk$Blank==1,]

#remove the species column and join with the full data set?
#photo.blnk$Species<-NULL

# remove the blank column
photo.blnk$Blank<-NULL

View(photo.blnk)
```

# rename the blank rate
```{r}
colnames(photo.blnk)[4]<-'blank.rate' 

write.csv(photo.blnk, 'Output/photo.blnk.csv') 
```

# join the blank data with the rest of the data
```{r}
Photo.R_full.blnk<-left_join(Photo.R_full, photo.blnk)
View(Photo.R_full.blnk)

```

# subtract the blanks
```{r}
Photo.R_full.blnk$umol.sec.corr<-Photo.R_full.blnk$umol.sec-Photo.R_full.blnk$blank.rate
```

# Normalize to organic biomass (ash free dry weight)
```{r}
#Calculate net P and R
Photo.R_full.blnk$umol.cm2.hr <- (Photo.R_full.blnk$umol.sec.corr*3600)/Photo.R_full.blnk$AFDW #mmol cm-2 hr-1

# remove NAs and blanks
Photo.R_full.blnk<-Photo.R_full.blnk[Photo.R_full.blnk$Blank==0,]

View(Photo.R_full.blnk)
```

#Calculate R
```{r}
#make ifelse statements to assign dark treatments as resp
Photo.R_full.blnk$rate.type <-ifelse(Photo.R_full.blnk$Light_Dark=='Light', "NP", "R")
Photo.R_full.blnk$rate.type<-as.factor(Photo.R_full.blnk$rate.type)

View(Photo.R_full.blnk)
```

#rename coral ID
```{r}
Photo.R_full.blnk$individual.ID <- str_split(Photo.R_full.blnk$ID,".{1}$", n = Inf, simplify = TRUE)[,1]
Photo.R_full.blnk$individual.ID <- as.factor(Photo.R_full.blnk$individual.ID)
View(Photo.R_full.blnk)

#make respiration positive
#Photo.R$umol.cm2.hr[Photo.R$PR=='Respiration']<-abs(Photo.R$umol.cm2.hr[Photo.R$PR=='Respiration'])
Photo.R_full.blnk$umol.cm2.hr<- -Photo.R_full.blnk$umol.cm2.hr

View(Photo.R_full.blnk)
```


```{r}
mydata <- read.csv('Output/Photo.R_full.csv')
View(mydata)
```

#######Raw data curve plots based on rate estimations#######
```{r}
#Combined mean curve plots across all species (Fig S1)

mydata <- read.csv('Output/Photo.R_full.csv')
library(ggplot2)

# Define the custom order of Species and their corresponding labels
species_order <- c("Pdarw", "Mplan", "Bgran", "Tcocc")
species_labels <- c("Bgran" = "B. grandis", "Mplan" = "M. plantaginea", "Pdarw" = "P. darwinii", "Tcocc" = "T. coccinea")

# Convert Species to factor with the custom order
mydata$Species <- factor(mydata$Species, levels = species_order)

# Specify custom colors for each species
species_colors <- c("Bgran" = "lemonchiffon2", "Pdarw" = "salmon",
                    "Mplan" = "mediumpurple1", "Tcocc" = "darkorange1")

# Modify the first plot
rawslope <- ggplot(mydata, aes(x = Temp.C, y = log.rate, group = Light_Dark, color = Light_Dark, fill = Species)) + 
  geom_smooth(se = TRUE, span = 1.5) +
  geom_point(size = 2, position = position_dodge(0.5)) +
  labs(y = expression(Rate~(log~mu~mol~O[2]~cm^-2~hr^-1)), x = "Temperature (ºC)") +
  theme_classic() +
  xlim(19, 38) + 
  scale_color_manual(values = c("Light" = species_colors["Bgran"], "Dark" = species_colors["Pdarw"])) +  # Use species_colors
  scale_fill_manual(values = species_colors, labels = species_labels) +  # Use species_colors for fill and set labels
  facet_wrap(~rate.type) +
  scale_x_continuous(labels = scales::comma_format(big.mark = "'", decimal.mark = "."), name = "Temperature (ºC)") +
  scale_y_continuous(labels = scales::comma_format(big.mark = "'", decimal.mark = "."), name = expression(Rate~(log~mu~mol~O[2]~cm^-2~hr^-1)))
rawslope

# Modify the second plot
FigS1 <- ggplot(mydata, aes(x = Temp.C, y = log.rate, col = Species, fill = Species)) +
  geom_point() + 
  geom_smooth(se = TRUE, span = 1.5) + 
  labs(y = expression(Rate~(log~mu~mol~O[2]~cm^-2~hr^-1)), x = "Temperature (ºC)") +
  theme_classic() +
  scale_color_manual(values = species_colors, labels = species_labels) +  # Use species_colors for color and set labels
  scale_fill_manual(values = species_colors, labels = species_labels) +  # Use species_colors for fill and set labels
  facet_wrap(~ Species * rate.type, 
             labeller = labeller(.multi_line = FALSE),
             scales = "free_y") +
  scale_x_continuous(labels = scales::comma_format(big.mark = "'", decimal.mark = "."), name = "Temperature (ºC)") +
  scale_y_continuous(labels = scales::comma_format(big.mark = "'", decimal.mark = "."), name = expression(Rate~(log~mu~mol~O[2]~cm^-2~hr^-1)))

FigS1

```
